Analysis of professional surfing tournaments with generalized Bradley-Terry

### 2021-08-10 inst/surfing_jsa.tex
rewording

Figure~\ref{piechartmax} shows the maximum likelihood estimate for the
23 competitors' strengths, which appears to show a wide range from
Florence at about 17.5\% down to Coffin at about 1.5\%.  The {\tt
hyper2} software may be used to formally assess one plausible null:
that the competitors all have equal skill, that is $H_0\colon
p_\mathrm{Andino} = p_\mathrm{Bourez}=\ldots=
p_\mathrm{Wright}=\frac{1}{23}$, and any differences in placings are
due to random variation.  The Method of Support \cite{edwards1973}
rejects the null (in favour of the evaluate) at $p=4\times 10^{-5}$.
There is thus strong evidence to suggest that the competitors are
indeed of different skills and the result is not random.  Note that
the competitors' points (that is, the final column of
table~\ref{resultstable}) does not allow one to come to this
conclusion as the points awarded are arbitrary.

However, points awarded may be compared with likelihood estimates:
Figure~\ref{compare_likelihood_points} shows a loose correlation
($R^2=0.64$; $p=3\times 10^{-6}$) and
Figure~\ref{compare_likelihood_points_rankings} shows how the rankings
differ when calculated by the points and likelihood systems.

>>>>

Figure~\ref{piechartmax} shows the maximum likelihood estimate for the
23 competitors' strengths, which appears to show a wide range from
Florence at about 17.5\% down to Coffin at about 1.5\%.  The {\tt
hyper2} software may be used to formally assess one plausible null:
that the competitors all have equal skill, that is $H_0\colon
p_\mathrm{Andino} = p_\mathrm{Bourez}=\ldots=
p_\mathrm{Wright}=\frac{1}{23}$, and any differences in placings are
due to random variation.  The Method of Support \cite{edwards1973}
rejects the null (in favour of the evaluate) at $p=4\times 10^{-5}$.
There is thus strong evidence to suggest that the competitors are
indeed of different skills and the result is not random.  Note that
the competitors' points (that is, the final column of
table~\ref{resultstable}) does not admit this type of analysis, as the
points awarded are arbitrary.

However, points awarded may be compared with likelihood estimates, and
one would expect high points totals to be associated with high
likelihoods.  Figure~\ref{compare_likelihood_points} shows a loose
correlation ($R^2=0.64$; $p=3\times 10^{-6}$) and
Figure~\ref{compare_likelihood_points_rankings} shows how the rankings
differ when calculated by the points and likelihood systems.  Note the
anomalous position of Florence, who did not compete in five venues due
to injury: this strongly affecting his points total but not his
likelihood-based ranking.
